---
title: "PCA_decathlon2"
author: "Zeyu CHEN"
date: "2018/11/19"
output: pdf_document
---

```{r , include}
library("FactoMineR")
library("factoextra")

data(decathlon2)
dim(decathlon2)

decathlon2.active <- decathlon2[1:23, 1:10]

#returns the first 4 lines
head(decathlon2.active[,1:6],4)

#In principal component analysis, variables are often scaled (i.e. standardized).

res.pca <- PCA(decathlon2.active, graph = FALSE)

print(res.pca)


```


# Eigenvalues / Variances
```{r , include}
#Extract the eigenvalues/variances of principal components
eig.val <- get_eigenvalue(res.pca)
eig.val

#Visualize the eigenvalues
fviz_eig(res.pca, addlabels = TRUE, ylim = c(0, 50))
```
# Graph of variables
```{r , include}
#Extract the results for variables
var <- get_pca_var(res.pca)

# Coordinates
var$coord

# Cos2: quality on the factore map
var$cos2

# Contributions to the principal components
head(var$contrib)
```

# Correlation circle
The correlation between a variable and a principal component (PC) is used as the coordinates of the variable on the PC. The representation of variables differs from the plot of the observations: The observations are represented by their projections, but the variables are represented by their correlations (Abdi and Williams 2010).
```{r , include}
head(var$coord, 4)
fviz_pca_var(res.pca, col.var = "black")
```

#Quality of representation
```{r , include}
head(var$cos2, 4)
library("corrplot")
corrplot(var$cos2, is.corr=FALSE)

# Total cos2 of variables on Dim.1 and Dim.2
fviz_cos2(res.pca, choice = "var", axes = 1:2)
```

It’s possible to color variables by their cos2 values using the argument col.var = "cos2". This produces a gradient colors. In this case, the argument gradient.cols can be used to provide a custom color. For instance, gradient.cols = c("white", "blue", "red") means that:

variables with low cos2 values will be colored in “white”
variables with mid cos2 values will be colored in “blue”
variables with high cos2 values will be colored in red
```{r , include}
# Color by cos2 values: quality on the factor map
fviz_pca_var(res.pca, col.var = "cos2",
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"), 
             repel = TRUE # Avoid text overlapping
             )

# Change the transparency by cos2 values
fviz_pca_var(res.pca, alpha.var = "cos2")
```

#Contributions of variables to PCs

```{r , include}
var$contrib
corrplot(var$contrib, is.corr=FALSE) 

# Contributions of variables to PC1
fviz_contrib(res.pca, choice = "var", axes = 1, top = 10)

# Contributions of variables to PC2
fviz_contrib(res.pca, choice = "var",axes= 2,top=10 )

# The total contribution to PC1 and PC2 is obtained with the following R code:
fviz_contrib(res.pca, choice = "var", axes = 1:2, top = 10)
```

