---
title: "Analysis and Description of data"
author: "Zeyu CHEN and Clément VEYSSIERE"
date: "2018/11/11"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# 1 The general nature of the problem

### Title of Database : Ailerons

### Sources: Experiments of Rui Camacho (rcamacho@garfield.fe.up.pt)
	
### Description :This data set addresses a control problem, namely flying a F16 aircraft. The attributes describe the status of the aeroplane, while the goal is to predict the control action on the ailerons of the aircraft

#2 The explanatory variables available
```{r , include=TRUE}

tab <- read.table("Ailerons/ailerons.data.txt",sep = ",")
tab <- data.frame(1:nrow(tab),tab,row.names = 1)

dim

```
###So we have 7154 instances and 41 variables,and each variable is a continuous number.

###here we will show you the fisrt ten rows of our data.
```{r , include=TRUE}
names(tab) = c("climbRate","Sgz", "p", "q", "curPitch", "curRoll", "absRoll", "diffClb", "diffRollRate", "diffDiffClb", "SeTime1", "SeTime2", "SeTime3", "SeTime4", "SeTime5", "SeTime6", "SeTime7", "SeTime8", "SeTime9", "SeTime10" , "SeTime11", "SeTime12", "SeTime13", "SeTime14", "diffSeTime1", "diffSeTime2", "diffSeTime3", "diffSeTime4", "diffSeTime5", "diffSeTime6", "diffSeTime7", "diffSeTime8", "diffSeTime9", "diffSeTime10", "diffSeTime11", "diffSeTime12", "diffSeTime13", "diffSeTime14", "alpha", "Se", "goal")
tab[1:5,]

tab$SeTime2 <- as.double(tab$SeTime2)
tab
summary(tab)

```
```{r, include = FALSE}
library(Amelia)
tab$SeTime2 = as.double(tab$SeTime2)
tab$diffSeTime2 = NULL


```

```{r, include = TRUE}
library(corrplot)

tabPredict <- tab
tab = as.data.frame(tab)
#tab = as.data.frame(scale(tab,center = FALSE))


sub <- sample(nrow(tab), 0.75 * nrow(tab))
tabTrain <- tab[sub,]
tabTest <- tab[-sub,]

res <- lm(goal~., data = tabTrain)
summary(res)

cor(tab)
par(cex=0.5)
corrplot(cor(tabTrain),type="lower")
```


```{r, include = TRUE}

library("lars")
library("glmnet")

#resboth <- step(res,direction='both')
reslasso<-  cv.glmnet(data.matrix(tabTrain[,-40]) ,y =as.vector(tabTrain$goal),type.measure="deviance",family="gaussian")
#summary(resboth)


#reslasso$lambda.min
reslasso<-  cv.glmnet(data.matrix(tabTrain[,-40]) ,y = as.vector(tabTrain$goal),type.measure="deviance",family="gaussian",alpha=1)


resrigde<-  cv.glmnet(data.matrix(tabTrain[,-40]) ,y =as.vector(tabTrain$goal),type.measure="deviance",family="gaussian",alpha=0)

resNet<-  cv.glmnet(data.matrix(tabTrain[,-40]) ,y =as.vector(tabTrain$goal),type.measure="deviance",family="gaussian",alpha=0.2)
#coef(reslasso$glmnet.fit, s = reslasso$lambda.min)
#coef(resrigde$glmnet.fit, s = resrigde$lambda.min)

pre <- predict(reslasso, newx = data.matrix(tabTest[,-40]), type="response" ,s=reslasso$lambda.min)  

preR <- predict(resrigde, newx = data.matrix(tabTest[,-40]), type="response" ,s=resrigde$lambda.min)  

preN <- predict(resNet, newx = data.matrix(tabTest[,-40]), type="response" ,s=resNet$lambda.min) 

Prelm <- predict(res,newdata = tabTest)
Preboth <- predict(resboth,newdata = tabTest)

sigma <- 3*sd(Prelm-tabTest$goal)
sigma
plot(Prelm,Prelm-tabTest$goal)
par(new=TRUE)
abline(h=sigma)




plot(Preboth,pre-tabTest$goal)

plot(pre,pre-tabTest$goal)


plot(preR,preR-tabTest$goal)

plot(preN,preN-tabTest$goal)
#sum squared residual
SSRlasso <- sum((pre-tabTest$goal)^2)
SSRrigde <- sum((preR-tabTest$goal)^2)
SSNet <- sum((preN-tabTest$goal)^2)
SSRlm <- sum((Prelm-tabTest$goal)^2)
SSRboth<- sum((Preboth-tabTest$goal)^2)


data.frame(SSRlm,SSRboth,SSRlasso,SSRrigde,SSNet)
```


```{r , include=TRUE}
#k-Nearest Neighbour Regressio
library(caret)

fit <- knnreg(x = tabTrain ,y = tabTrain$goal, k = 5) 

pred <- predict(fit, tabTest)

plot(pred,pred-tabTest$goal)

sum((pred-tabTest$goal)^2)


```




```{r , include=TRUE}

kemans_lm <- function(k)
{
  i<-0
  kmTrain <- kmeans(tabTrain[,-40], k)
  tabls <-list()
  tabT <- data.frame()
  cl <- rep()
  for(i in 1:k)
  {
     tabls[[i]] <- tabTrain[kmTrain$cluster==i,]
     tabT <- rbind(tabT,tabls[[i]])
     cl <- factor(c(cl,rep(as.character(i),dim(tabls[[i]])[1])))
  }
  
  class <- knn(tabT, tabTest, cl, k = k)
  RSS<-0
  i<-0
  for(i in 1:k)
  {
    Test <- tabTest[class==i,]
    resls <- lm(goal~.,data = tabls[[i]])
    resboth <- step(resls,direction='both',trace = 0)
    Preboth <- predict(resboth ,newdata = Test)
    RSS <- RSS+ sum((Preboth - Test$goal)^2)
  }
  return(RSS)
}

resls <-1
i <- 0
for(i in 1:3)
{
  rss <- kemans_lm(i)
  if(rss < resls)
  {
    index <- i 
    resls <- rss
  }
}
resls 
index
```




```{r , include=TRUE}

kmTrain <- kmeans(tabTrain, 2)

tab1 <- tabTrain[kmTrain$cluster==1,]
tab2 <- tabTrain[kmTrain$cluster==2,]
#tab3 <- tabTrain[kmTrain$cluster==3,]
#tab4<- tabTrain[kmTrain$cluster==4,]
#tab5 <- tabTrain[kmTrain$cluster==5,]

tabT <- rbind(tab1,tab2)
#cl <- factor(c(rep("1",dim(tab1)[1]),rep("2",dim(tab2)[1]),rep("3",dim(tab3)[1]),rep("4",dim(tab4)[1]),rep("5",dim(tab5)[1])))
cl <- factor(c(rep("1",dim(tab1)[1]),rep("2",dim(tab2)[1])))
k <- knn(tabT, tabTest, cl, k = 2)

tabTest1 <- tabTest[k==1,]
tabTest2 <- tabTest[k==2,]
#tabTest3 <- tabTest[k==3,]
#tabTest4 <- tabTest[k==4,]
#tabTest5 <- tabTest[k==5,]

res1 <- lm(goal~.,data = tab1)
res2 <- lm(goal~.,data = tab2)
#res3<- lm(goal~.,data = tab3)
#res4<- lm(goal~.,data = tab4)
#res5<- lm(goal~.,data = tab5)

resboth1 <- step(res1,direction='both')
Preboth1 <- predict(resboth1,newdata = tabTest1)

resboth2 <- step(res2,direction='both')
Preboth2 <- predict(resboth2,newdata = tabTest2)

#resboth3 <- step(res3,direction='both')
#Preboth3 <- predict(resboth3,newdata = tabTest3)

#resboth4 <- step(res4,direction='both')
#Preboth4 <- predict(resboth4,newdata = tabTest4)

#resboth5 <- step(res5,direction='both')
#Preboth5 <- predict(resboth5,newdata = tabTest5)

sum((Preboth1-tabTest1$goal)^2)+sum((Preboth2-tabTest2$goal)^2)
#sum((Preboth1-tabTest1$goal)^2)+sum((Preboth2-tabTest2$goal)^2)+sum((Preboth3-tabTest3$goal)^2)+sum((Preboth4-tabTest4$goal)^2)+sum((Preboth5-tabTest5$goal)^2)

```
### We can find that all the points are roughly symmetrically distributed around the x-axis,so it will be reasonable to use linear regression.








